The Effects of Racial Identity and Discrimination on Political Participation and Political Attitudes

### 2024-12-15 draft_finalreport.qmd
updating draft of final

---
title: "The Effects of Racial Identity and Discrimination on Political Participation and Political Attitudes"
author: "Peter de Guzman, Jennifer Li, Nakiyah Dhariwala, Sizhe Chen"
format: html
date: 2024-12-08
editor: visual
---

@@

```{r echo=FALSE, results=FALSE}
# Remove all redundant columns
redundant_columns <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", 
                       "Q629R5", "Q629R6", "Q629R7",
                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")

@@

```{r echo=FALSE, results=FALSE}
# Rename remaining columns in dataset_clean
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
  "Gender",                                               # S3B
  "TreatedUnfairlyOrDiscriminatedAgainst",                # Q627
  "DiscriminationImpactOnLifeSatisfaction",              # Q633
  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
  "Education",                                           # educ
  "SexOrientation",                                      # sex_orientation
  "Rural",                                               # rural
  "Race",                                                # race_recoded
  "Age"                                                  # age_recoded
)

@@

# ordinal model

# ```{r, echo=FALSE, results=FALSE}
# # data load and cleaning for ordinal regression model
# required_cols_ord <- c("Q29", # Outcome Variable; interest in politics
#                    "Q271", # Predictor Variable; importance of race to identity
#                    "S2_RACE_PRIME", # Race
#                    "S3", # Orientation
#                    "S3B", # Gender
#                    "S5_AGE", # Age Category
#                    "S13", # Educational Attainment
#                    "S14", # Community Type
#                    "S7", # Nativity
#                    "Q21", # Party ID
#                    "S2_M", # Multiracial
#                    
#                    # Religion-related variables
#                    "Q58R1",            # Catholic
#                    "Q58R2",            # Protestant
#                    "Q58R3",            # Other Christian
#                    "Q58R4",            # Jewish
#                    "Q58R5",            # Muslim
#                    "Q58R6",            # Hindu
#                    "Q58R7",            # Buddhist
#                    "Q58R8",            # Mormon/LDS
#                    "Q58R9",            # Folk Religion
#                    "Q58R10",           # Daoist
#                    "Q58R11",           # Shamanism
#                    "Q58R12",           # Ancestral Religion
#                    "Q58R13",           # Spiritual but not religious
#                    "Q58R14",           # Atheist
#                    "Q58R15",            # Agnostic
#                    
#                    # Variables capturing experiences with racism
#                    "Q627", # Treated unfairly or discriminated against ,
#                    "Q628", # Location of discrimination
#                    "Q629R1" # Unfair treatment due to racial background/ethnicity
#                    )
#                  
#                    
#                    
# # Create new df
# data_filtered <- dataset[, required_cols_ord]
# 
# # Convert the required columns to factors
# data_filtered[required_cols] <- lapply(data_filtered[required_cols_ord], as.factor)
# 
# # Convert individual religion columns to numeric
# for (i in 1:15) {
#   col_name <- paste0("Q58R", i, "_numeric")
#   data_filtered[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data_filtered[[paste0("Q58R", i)]]))
# }
# 
# # Create a new Religion variable
# data_filtered$Religion <- "None"
# data_filtered$Religion[data_filtered$Q58R1_numeric == 1] <- "Catholic"
# data_filtered$Religion[data_filtered$Q58R2_numeric == 1] <- "Protestant"
# data_filtered$Religion[data_filtered$Q58R3_numeric == 1] <- "Other Christian"
# data_filtered$Religion[data_filtered$Q58R4_numeric == 1] <- "Jewish"
# data_filtered$Religion[data_filtered$Q58R5_numeric == 1] <- "Muslim"
# data_filtered$Religion[data_filtered$Q58R6_numeric == 1] <- "Hindu"
# data_filtered$Religion[data_filtered$Q58R7_numeric == 1] <- "Buddhist"
# data_filtered$Religion[data_filtered$Q58R8_numeric == 1] <- "Mormon/LDS"
# data_filtered$Religion[data_filtered$Q58R9_numeric == 1] <- "Folk Religion"
# data_filtered$Religion[data_filtered$Q58R10_numeric == 1] <- "Daoist"
# data_filtered$Religion[data_filtered$Q58R11_numeric == 1] <- "Shamanism"
# data_filtered$Religion[data_filtered$Q58R12_numeric == 1] <- "Ancestral Religion"
# data_filtered$Religion[data_filtered$Q58R13_numeric == 1] <- "Spiritual but not religious"
# data_filtered$Religion[data_filtered$Q58R14_numeric == 1] <- "Atheist"
# data_filtered$Religion[data_filtered$Q58R15_numeric == 1] <- "Agnostic"
# 
# # Convert Religion to a factor
# data_filtered$Religion <- factor(data_filtered$Religion,
#                                levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
#                                           "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
#                                           "Daoist", "Shamanism", "Ancestral Religion",
#                                           "Spiritual but not religious", "Atheist", "Agnostic", "None"))
# 
# # Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
# redundant_cols <- c(
#   "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
#   "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
#   "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
#   "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
#   "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
#   "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
# )
# 
# # Remove redundant columns from the dataset
# dataset_clean_ord <- data_filtered[, !(names(data_filtered) %in% redundant_cols)]
# 
# dataset_clean_ord <- dataset_clean_ord %>% mutate(educ = case_when(S13 == "(1) Grades 1-8" ~ "No college experience",
#                                                            S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
#                                                            S13 == "(3) High School graduate or GED" ~ "No college experience",
#                                                            S13 == "(4) Some college" ~ "Some college experience",
#                                                            S13 == "(5) Associates, 2-year degree" ~ "College graduate",
#                                                            S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
#                                                            S13 == "(7) Post-graduate degree" ~ "College graduate"),
#                                           sex_orientation = case_when(S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
#                                                                      S3 == "(02) Gay or lesbian" ~ "Queer",
#                                                                      S3 == "(03) Bisexual" ~ "Queer",
#                                                                      S3 == "(04) Something else (Specify)" ~ "Queer",
#                                                                      S3 == "(88) Refused" ~ "Refused/Don't Know",
#                                                                      S3 == "(99) I don" ~ "Refused/Don't Know"),
#                                           rural = case_when(S14 == "(1) Large urban area" ~ "Urban/suburban",
#                                                             S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
#                                                             S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
#                                                             S14 == "(4) Small town or small city" ~ "Urban/suburban",
#                                                             S14 == "(5) Rural area" ~ "Rural"),
#                                           white = case_when(S2_RACE_PRIME == "(1) White" ~ "White", 
#                                                             S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Non-White",
#                                                             S2_RACE_PRIME == "(3) Black or African American" ~ "Non-White",
#                                                             S2_RACE_PRIME == "(4) Asian American" ~ "Non-White",
#                                                             S2_RACE_PRIME == "(5) American Indian/Native American" ~ "Non-White",
#                                                             S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Non-White",
#                                                             S2_RACE_PRIME == "(7) Native Hawaiian" ~ "Non-White",
#                                                             S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "Non-White"),
#                                           race_recoded = case_when(S2_RACE_PRIME == "(1) White" ~ "White",
#                                                                    S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
#                                                                    S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
#                                                                    S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
#                                                                    S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
#                                                                    S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
#                                                                    S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
#                                                                    S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"),
#                                           age_recoded = case_when(S5_AGE == "(1) 16-17" ~ "16-17",
#                                                                   S5_AGE == "(2) 18-29" ~ "18-29",
#                                                                   S5_AGE == "(3) 30-39" ~ "30-39",
#                                                                   S5_AGE == "(4) 40-49" ~ "40-49",
#                                                                   S5_AGE == "(5) 50-59" ~ "50-59",
#                                                                   S5_AGE == "(6) 60-69" ~ "60-69",
#                                                                   S5_AGE == "(7) 70 +" ~ "70+"))
# 
# dataset_clean_ord <- dataset_clean_ord %>% mutate(race_recoded = factor(dataset_clean_ord$race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
#                                          age_recoded = factor(dataset_clean_ord$age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
#                                          educ = factor(dataset_clean_ord$educ, levels = c("No college experience", "Some college experience", "College graduate")),
#                                          white = factor(dataset_clean_ord$white, levels = c("White", "Non-White")),
#                                          rural = factor(dataset_clean_ord$rural, levels= c("Rural", "Urban/suburban")),
#                                          sex_orientation = factor(dataset_clean_ord$sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know")))
# 
# 
# ```

@@

The model also included the predictor variable regarding the importance of belonging to a racial group to one’s identity, a categorical variable that captured responses to the question: “How important is being [RACIAL GROUP] to your identity?”, with response options “Extremely important”, “Very important”, “Moderately important”, “Slightly important”, and “Not at all important”. The respondent’s identified racial group was brought into the survey question based on their prior survey responses. Over half (56%) of respondents reported that being a member of their racial group was very or extremely important to their identity. Only 23% of respondents said that being a member of their racial group was slightly or not at all important to their identity.

@@

(Insert Odds Ratio Plot for RQ1)

@@

@@

To examine factors influencing the importance of anti-discrimination policies, the model included predictors like personal experiences with racism (measured via ‘Unfair treatment or discrimination’ and its impact on life satisfaction), alongside demographic variables such as primary race, sexual orientation, gender, age, education, and community type. Variables Q629R1–Q629R7, which captured specific experiences of unfair treatment (e.g., due to race or gender), were excluded due to substantial missing data (>52%). However, this was justified since the broader question on ‘unfair treatment or discrimination’ retained the key essence of experiences with racism.

@@

```{r}

@@

The balanced accuracy was 55.26%, reflecting the model’s challenges in handling class imbalance. Despite achieving statistical significance with McNemar’s test (p-value < 2e-16), the model’s predictions deviate considerably from the true data.
In conclusion, while the model effectively identifies respondents who do not prioritize anti-discrimination policies, it performs poorly in predicting those who do. This imbalance signals the need for improvement in predicting the minority class.

@@

Following the resource model of participation, age and increases in educational attainment appeared to be associated with higher odds of reporting a higher level of interest in politics. The most significant positive predictors of increased levels of interest in politics were an individual’s age and level of educational attainment. In particular, individuals who graduated from college had much higher odds of reporting a higher level of interest in politics than individuals with no college experience. This follows popular findings that educational attainment can lead to increases in political participation [8], and that older Americans vote at higher rates than Americans ages 18-29 years old [9].

@@

Despite the value of this survey’s oversamples of racial and ethnic minority communities, some limitations were present in our study. For the ordinal logistic regression, we used educational attainment as a proxy for socioeconomic status. This is a commonly used proxy measure in behavioral and social sciences [10]. Further research on the influence of socioeconomic status of political participation for racial/ethnic minorities and the use of other measures such as household income or wealth would help us understand if the aforementioned resource model of participation is indeed generalizable to other groups in the American electorate. 
The effect of race on political interest appeared to be more pronounced for members of very small racial/ethnic minority groups in our analysis such as Asian Americans, Native Americans, Native Hawaiians and Pacific Islanders. Both models had low specificity, suggesting that they struggled to correctly identify cases where respondents had low levels of political interest or respondents who considered anti-discrimination policies to be important. Similarly, in our second research question, while the model effectively identified respondents who do not prioritize anti-discrimination policies, it struggled to predict those who do. This imbalance underscores the challenges of accurately modeling the minority class—respondents who view anti-discrimination policies as important—and highlights the need for more nuanced approaches to capture these perspectives.

@@

@@

1.    Noe-Bustamante, A. B., Christine Tamir, Lauren Mora and Luis. (2020, August 20). Facts on U.S. immigrants, 2018. Pew Research Center. https://www.pewresearch.org/race-and-ethnicity/2020/08/20/facts-on-u-s-immigrants/
2.    Jones, N., Marks, R., Ramirez, R., & Rios-Vargas, M. (2021, August 12). Improved Race and Ethnicity Measures Reveal U.S. Population Is Much More Multiracial. The United States Census Bureau. https://www.census.gov/library/stories/2021/08/improved-race-ethnicity-measures-reveal-united-states-population-much-more-multiracial.html
3.    Brady, H. E., Verba, S., & Schlozman, K. L. (1995). Beyond Ses: A Resource Model of Political Participation. The American Political Science Review, 89(2), 271–294. https://doi.org/10.2307/2082425
4.    Leighley, J. E. (2021). Strength in Numbers? Princeton University Press.
5.    Besco, R., Garcia-Rios, S., Lagodny, J., Lajevardi, N., Oskooii, K., & Tolley, E. (2022). Fight not flight: The effects of explicit racism on minority political engagement. Electoral Studies, 80, 102515. https://doi.org/10.1016/j.electstud.2022.102515
6.    Wood, Z. J., & John-Henderson, N. A. (2024). Perceived discrimination, political efficacy, and political participation in American Indian adults. Frontiers in Political Science, 6. https://doi.org/10.3389/fpos.2024.1328521
7.    Frasure, Lorrie, Wong, Janelle, Vargas, Edward, and Barreto, Matt. Collaborative Multi-racial Post-election Survey (CMPS), United States, 2020. Inter-university Consortium for Political and Social Research [distributor], 2024-06-11. https://doi.org/10.3886/ICPSR39096.v1
8.    Kim, Y. (2023). Absolutely Relative: How Education Shapes Voter Turnout in the United States. Social Indicators Research, 168(1-3), 447–469. https://doi.org/10.1007/s11205-023-03146-1
9.    McDonald, M. (n.d.). Voter Turnout Demographics [Review of Voter Turnout Demographics]. U.S. Elections Project. Retrieved December 13, 2024, from https://www.electproject.org/election-data/voter-turnout-demographics
10.   Oakes, M. (n.d.). Measuring Socioeconomic Status 1. Learning Objectives. https://obssr.od.nih.gov/sites/obssr/files/Measuring-Socioeconomic-Status.pdf

@@

### Model Summary - OLR Model with Interactions
```{r}
dataset_clean_allvars$Q271_reordered <- factor(dataset_clean_allvars$Q271,
                            levels = c("(5) Not at all important",
                                     "(4) Slightly important",
                                     "(3) Moderately important",
                                     "(2) Very important",
                                     "(1) Extremely important"))
dataset_clean_allvars$Q29_reordered <- factor(dataset_clean_allvars$Q29,
                            levels = c("(4) Not at all interested in politics",
                                     "(3) Not that interested in politics",
                                     "(2) Somewhat interested",
                                     "(1) Very interested in politics"), ordered=TRUE)

confusionMatrix(predict(ord_mod_interactions), dataset_clean$Q29, mode="everything")

>>>>

---
title: "The Effects of Racial Identity and Discrimination on Political Participation and Political Attitudes"
author: "Peter de Guzman, Jennifer Li, Nakiyah Dhariwala, Sizhe Chen"
format: pdf
date: 2024-12-08
editor: visual
---

@@

```{r echo=FALSE, results=FALSE}
# Remove all redundant columns
redundant_columns <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", 
                       "Q629R5", "Q629R6", "Q629R7",
                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14", "Q29", "Q21")

@@

```{r echo=FALSE, results=FALSE}
# Rename remaining columns in dataset_clean
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
  "RacialBelonging",                                      #Q271
  "Gender",                                               # S3B
  "TreatedUnfairlyOrDiscriminatedAgainst",                # Q627
  "DiscriminationImpactOnLifeSatisfaction",              # Q633
  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
  "Education",                                           # educ
  "SexOrientation",                                      # sex_orientation
  "Rural",                                               # rural
  "Race",                                                # race_recoded
  "Age"                                                  # age_recoded
)

@@

@@

The model also included the predictor variable regarding the importance of belonging to a racial group to one’s identity, a categorical variable that captured responses to the question: “How important is being \[RACIAL GROUP\] to your identity?”, with response options “Extremely important”, “Very important”, “Moderately important”, “Slightly important”, and “Not at all important”. The respondent’s identified racial group was brought into the survey question based on their prior survey responses. Over half (56%) of respondents reported that being a member of their racial group was very or extremely important to their identity. Only 23% of respondents said that being a member of their racial group was slightly or not at all important to their identity.

@@

```{r, echo=FALSE, results=FALSE}
dataset_clean_allvars$Q271_reordered <- factor(dataset_clean_allvars$Q271,
                            levels = c("(5) Not at all important",
                                     "(4) Slightly important",
                                     "(3) Moderately important",
                                     "(2) Very important",
                                     "(1) Extremely important"))
dataset_clean_allvars$Q29_reordered <- factor(dataset_clean_allvars$Q29,
                            levels = c("(4) Not at all interested in politics",
                                     "(3) Not that interested in politics",
                                     "(2) Somewhat interested",
                                     "(1) Very interested in politics"), ordered=TRUE)

confusionMatrix(predict(ord_mod_interactions), dataset_clean_allvars$Q29, mode="everything")

brant(ord_mod_interactions)

```{r, echo=FALSE}
odds_ratio_plot_labelled <- function(model) {
  # Get odds ratios and CIs
  or_data <- exp(cbind(OR=coef(model), confint(model)))

  # Calculate p-values
  p_values <- pnorm(-abs(summary(model)$coefficients[,"t value"]))*2

  # Keep only significant variables
  sig_indices <- p_values[names(coef(model))] < 0.05
  or_data_sig <- or_data[sig_indices, ]

  # Create a mapping for variable names with grouping prefixes
  rename_map <- c(
    # Education Group
    "educCollege graduate" = "1. Education: College Graduate",
    "educSome college experience" = "1. Education: Some College",

    # Gender Group
    "S3B(2) Woman" = "2. Gender: Woman",
    "S3B(3) Non-binary" = "2. Gender: Non-binary",

    # Age Group
    "age_recoded70+" = "3. Age: 70+",
    "age_recoded60-69" = "3. Age: 60-69",
    "age_recoded50-59" = "3. Age: 50-59",

    # Race Group
    "race_recodedBlack" = "4. Race: Black",
    "race_recodedAsian American" = "4. Race: Asian",
    "race_recodedLatino" = "4. Race: Latino",
    "race_recodedAIAN" = "4. Race: Native",
    "race_recodedNHPI" = "4. Race: NHPI",

    # Racial Belonging Group
    "Q271_reordered(2) Very important" = "5. Racial Belonging: Very Important",
    "Q271_reordered(1) Extremely important" = "5. Racial Belonging: Extremely Important",

    # Party ID Group
    "Q21(2) Democrat" = "6. Party ID: Democrat",
    "Q21(3) Independent" = "6. Party ID: Independent",
    "Q21(4) Other party" = "6. Party ID: Other",

    # Interactions
    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "7. Interaction: Racial Belonging & Gender: Woman"
  )

  # Create a data frame for plotting
  plot_data <- data.frame(
    name = rename_map[rownames(or_data_sig)],
    or = or_data_sig[,1],
    ci_lower = or_data_sig[,2],
    ci_upper = or_data_sig[,3],
    stringsAsFactors = FALSE
  )

  # Custom sorting function
  sort_by_group_and_name <- function(x) {
    groups <- as.numeric(sub("^(\\d+)\\..+$", "\\1", x))
    names <- sub("^\\d+\\.[^:]+: (.+)$", "\\1", x)
    sort_df <- data.frame(
      original = x,
      group = groups,
      name = names,
      stringsAsFactors = FALSE
    )
    sort_df <- sort_df[order(sort_df$group, -xtfrm(sort_df$name)), ]
    return(sort_df$original)
  }

  # Sort the data
  plot_data$name <- factor(plot_data$name,
                          levels = rev(sort_by_group_and_name(plot_data$name)))

  # Create plot
  ggplot(plot_data, aes(y = name)) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
    geom_point(aes(x = or), size = 4, color = "blue") +
    geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0.3, color = "black") +
    theme_minimal() +
    theme(
      axis.text.y = element_text(size = 6, color = "black"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      axis.title.x = element_text(size = 12, face = "bold"),
      axis.title.y = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.margin = margin(1, 1, 1, 1, "cm"),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.minor.y = element_blank()
    ) +
    labs(
      title = "Significant Variables with Confidence Intervals",
      x = "Odds Ratio",
      y = "Variables"
    )
}

# Generate the plot
odds_ratio_plot_labelled(ord_mod_interactions)

@@

| Variable Group | Reference Level |
|------------------------------------|------------------------------------|
| 1\. Education | No college experience |
| 2\. Gender | Man |
| 3\. Age | 18-29 years old |
| 4\. Race | White |
| 5\. Racial Belonging | Not At All Important |
| 6\. Party ID | Republican |
| 7\. Interaction: Racial Belonging & Gender: Woman | Man, Not At All Important |

@@

To examine factors influencing the importance of anti-discrimination policies, the model included predictors like personal experiences with racism (measured via ‘Unfair treatment or discrimination’ and its impact on life satisfaction), alongside demographic variables such as primary race, sexual orientation, gender, age, education, and community type. Variables Q629R1–Q629R7, which captured specific experiences of unfair treatment (e.g., due to race or gender), were excluded due to substantial missing data (\>52%). However, this was justified since the broader question on ‘unfair treatment or discrimination’ retained the key essence of experiences with racism.

@@

```{r, echo=FALSE}

@@

The balanced accuracy was 55.26%, reflecting the model’s challenges in handling class imbalance. Despite achieving statistical significance with McNemar’s test (p-value \< 2e-16), the model’s predictions deviate considerably from the true data. In conclusion, while the model effectively identifies respondents who do not prioritize anti-discrimination policies, it performs poorly in predicting those who do. This imbalance signals the need for improvement in predicting the minority class.

@@

Following the resource model of participation, age and increases in educational attainment appeared to be associated with higher odds of reporting a higher level of interest in politics. The most significant positive predictors of increased levels of interest in politics were an individual’s age and level of educational attainment. In particular, individuals who graduated from college had much higher odds of reporting a higher level of interest in politics than individuals with no college experience. This follows popular findings that educational attainment can lead to increases in political participation \[8\], and that older Americans vote at higher rates than Americans ages 18-29 years old \[9\].

@@

Despite the value of this survey’s oversamples of racial and ethnic minority communities, some limitations were present in our study. For the ordinal logistic regression, we used educational attainment as a proxy for socioeconomic status. This is a commonly used proxy measure in behavioral and social sciences \[10\]. Further research on the influence of socioeconomic status of political participation for racial/ethnic minorities and the use of other measures such as household income or wealth would help us understand if the aforementioned resource model of participation is indeed generalizable to other groups in the American electorate. The effect of race on political interest appeared to be more pronounced for members of very small racial/ethnic minority groups in our analysis such as Asian Americans, Native Americans, Native Hawaiians and Pacific Islanders. Both models had low specificity, suggesting that they struggled to correctly identify cases where respondents had low levels of political interest or respondents who considered anti-discrimination policies to be important. Similarly, in our second research question, while the model effectively identified respondents who do not prioritize anti-discrimination policies, it struggled to predict those who do. This imbalance underscores the challenges of accurately modeling the minority class—respondents who view anti-discrimination policies as important—and highlights the need for more nuanced approaches to capture these perspectives.

@@

\newpage

@@

1.  Noe-Bustamante, A. B., Christine Tamir, Lauren Mora and Luis. (2020, August 20). Facts on U.S. immigrants, 2018. Pew Research Center. https://www.pewresearch.org/race-and-ethnicity/2020/08/20/facts-on-u-s-immigrants/
2.  Jones, N., Marks, R., Ramirez, R., & Rios-Vargas, M. (2021, August 12). Improved Race and Ethnicity Measures Reveal U.S. Population Is Much More Multiracial. The United States Census Bureau. https://www.census.gov/library/stories/2021/08/improved-race-ethnicity-measures-reveal-united-states-population-much-more-multiracial.html
3.  Brady, H. E., Verba, S., & Schlozman, K. L. (1995). Beyond Ses: A Resource Model of Political Participation. The American Political Science Review, 89(2), 271–294. https://doi.org/10.2307/2082425
4.  Leighley, J. E. (2021). Strength in Numbers? Princeton University Press.
5.  Besco, R., Garcia-Rios, S., Lagodny, J., Lajevardi, N., Oskooii, K., & Tolley, E. (2022). Fight not flight: The effects of explicit racism on minority political engagement. Electoral Studies, 80, 102515. https://doi.org/10.1016/j.electstud.2022.102515
6.  Wood, Z. J., & John-Henderson, N. A. (2024). Perceived discrimination, political efficacy, and political participation in American Indian adults. Frontiers in Political Science, 6. https://doi.org/10.3389/fpos.2024.1328521
7.  Frasure, Lorrie, Wong, Janelle, Vargas, Edward, and Barreto, Matt. Collaborative Multi-racial Post-election Survey (CMPS), United States, 2020. Inter-university Consortium for Political and Social Research \[distributor\], 2024-06-11. https://doi.org/10.3886/ICPSR39096.v1
8.  Kim, Y. (2023). Absolutely Relative: How Education Shapes Voter Turnout in the United States. Social Indicators Research, 168(1-3), 447–469. https://doi.org/10.1007/s11205-023-03146-1
9.  McDonald, M. (n.d.). Voter Turnout Demographics \[Review of Voter Turnout Demographics\]. U.S. Elections Project. Retrieved December 13, 2024, from https://www.electproject.org/election-data/voter-turnout-demographics
10. Oakes, M. (n.d.). Measuring Socioeconomic Status 1. Learning Objectives. https://obssr.od.nih.gov/sites/obssr/files/Measuring-Socioeconomic-Status.pdf

\newpage

@@

### Ordinal Model Overall Statistics

|                           | Metric           |
|---------------------------|------------------|
| Accuracy                  | 0.4503           |
| 95% CI                    | (0.4429, 0.4577) |
| No Information Rate (NIR) | 0.4339           |
| P Value \[Acc \> NIR\]    | 6.105e-06        |
| Kappa                     | 0.0787           |
| Mcnemar's Test P-Value    | \<2.2e-16        |

### Confusion Matrix

|   | **Actual - Very interested in politics** | **Actual -Somewhat interested** | **Actual - Not that interested in politics** | **Actual - Not at all interested in politics** |
|---------------|---------------|---------------|---------------|---------------|
| **Predicted - Very interested in politics** | 650 | 505 | 120 | 31 |
| **Predicted - Somewhat interested** | 3567 | 6830 | 3202 | 1550 |
| **Predicted - Not that interested in politics** | 57 | 181 | 203 | 186 |
| **Predicted - Not at all interested in politics** | 32 | 96 | 118 | 217 |

### Statistics by Class:

|   | Very interested in politics | Somewhat interested | Not that interested in politics | Not at all interested in politics |
|---------------|---------------|---------------|---------------|---------------|
| Sensitivity | 0.15095 | 0.8973 | 0.05572 | 0.10938 |
| Specificity | 0.95045 | 0.1625 | 0.96950 | 0.98419 |
| Balanced Accuracy | 0.55070 | 0.5299 | 0.51261 | 0.54678 |

| Variable | Odds Ratio | 2.5% (Confidence Interval) | 97.5% (Confidence Interval) |
|------------------|------------------|------------------|------------------|
| Q271 (4) Slightly important | 0.9672589 | 0.811845068 | 1.1525850 |
| Q271 (3) Moderately important | 0.9776608 | 0.845556718 | 1.1303745 |
| Q271 (2) Very important | 1.1952004 | 1.033019394 | 1.3828451 |
| Q271 (1) Extremely important | 1.6474969 | 1.423670208 | 1.9066091 |
| S3B(2) Woman | 0.6294256 | 0.540015144 | 0.7336072 |
| S3B(3) Non-binary | 1.8576658 | 1.002509318 | 3.4617409 |
| S3B(4) Something else (Specify) | 2.1359493 | 0.377740111 | 13.5124525 |
| race Black | 0.8676620 | 0.790376010 | 0.9524806 |
| race Asian American | 0.5945609 | 0.541964870 | 0.6522143 |
| race Latino | 0.7946622 | 0.575235415 | 0.7848896 |
| race AIAN | 0.6719204 | 1.27404185 | 1.7384045 |
| race NHPI | 0.5606006 | 0.458205702 | 0.6860792 |
| race Arab/MENA | 1.1589043 | 0.899447122 | 1.4956228 |
| age 30-39 | 1.0310801 | 0.946696738 | 1.0563119 |
| age 40-49 | 0.9247623 | 0.98549367 | 1.1229868 |
| age 50-59 | 1.0979487 | 1.000027913 | 1.2054892 |
| age 60-69 | 1.3238874 | 1.205240800 | 1.4542874 |
| age 70+ | 1.6896464 | 1.510226543 | 1.8907050 |
| educSome college experience | 1.5264694 | 1.402573135 | 1.6614197 |
| educCollege graduate | 2.2337286 | 2.072162224 | 2.4081069 |
| Q21(2) Democrat | 1.1333513 | 1.045765624 | 1.2282275 |
| Q21(3) Independent | 0.5806653 | 0.533801211 | 0.6315870 |
| Q21(4) Other party | 0.2127592 | 0.182654104 | 0.2477198 |
| Q271 4) Slightly important:S3B(2) Woman | 0.9955364 | 0.782564309 | 1.2663323 |
| Q271 (3) Moderately important:S3B(2) Woman | 0.9461281 | 0.780040309 | 1.1475378 |
| Q271 (2) Very important:S3B(2) Woman | 0.9791675 | 0.810674379 | 1.1826356 |
| Q271 (1) Extremely important:S3B(2) Woman | 0.8114102 | 0.674275440 | 0.9763646 |
| Q271 (4) Slightly important:S3B(3) Non-binary | 0.6505015 | 0.198565161 | 2.1505278 |
| Q271 (3) Moderately important:S3B(3) Non-binary | 0.6563254 | 0.249594245 | 1.7277048 |
| Q271 (2) Very important:S3B(3) Non-binary | 0.5140828 | 0.205132143 | 1.2905898 |
| Q271 (1) Extremely important:S3B(3) Non-binary | 0.8770903 | 0.354410521 | 2.1835545 |
| Q271 (4) Slightly important:S3B(4) Something else (Specify) | 0.2279084 | 0.023857021 | 2.0164000 |
| Q271 (3) Moderately important:S3B(4) Something else (Specify) | 1.7158304 | 0.146865519 | 19.7516328 |
| Q271 2) Very important:S3B(4) Something else (Specify) | 0.1049752 | 0.006213617 | 1.4210951 |
| Q271 1) Extremely important:S3B(4) Something else (Specify) | 0.6584933 | 0.079780177 | 4.9850479 |
